### 2017-08-16 glossary/acronyms.tex
Revise intro & add omnivore's dilemma citation

\newacronym{caf}{CAF}{coarray Fortran}
\newacronym{mini-app}{mini-app}{mini-application}
\newacronym{ncar}{NCAR}{National Center for Atmospheric Research}
\newacronym{wrf-hydro}{WRF-Hydro}{Weather Research and Forcecasting -- Hydrological Model}
\newacronym{spmd}{SPMD}{Single-Program Multiple-Data}
\newacronym{url}{URL}{Uniform Resource Locator}
\newacronym{tdd}{TDD}{test-driven development}
\newacronym{upc++}{UPC++}{Unified Parallel C++}
\newacronym{upc}{UPC}{Unified Parallel C}
\newacronym{lbnl}{LBNL}{Lawrence Berkeley National Laboratoory}
\newacronym{ide}{IDE}{integrated development environment}
\newacronym{html}{HTML}{Hypertext Markup Language}
\newacronym{icar}{ICAR}{The Intermediate Complexity Atmospheric Research}
\newacronym{exaflops}{exaFLOPS}{$10^{18}$ floating-point operations per second}
\newacronym{hpc}{HPC}{high-performance computing}
\newacronym{nasa}{NASA}{National Aeronautics and Space Administration}
\newacronym{nersc}{NERSC}{National Energy Research Scientific Computing Center}
\newacronym{doe}{DOE}{Department of Energy}
\newacronym{us}{US}{United States}
\newacronym{oop}{OOP}{Object-Oriented Programming}
\newacronym{ood}{OOD}{Object-Oriented Design}
\newacronym{uml}{UML}{Unified Modeling Language}
\newacronym{psblas}{PSBLAS}{Parallel Sparse Basic Linear Algebra Subroutines}
\newacronym{gpu}{GPU}{Graphics Processing Unit}
\newacronym{gpgpu}{GPGPU}{General-Purpose Graphics Processing Unit}
\newacronym{simt}{SIMT}{Single-Instruction, Multiple-Threads}
\newacronym{simd}{SIMD}{Single-Instruction, Multiple-Data}
\newacronym{2d}{2D}{Two-Dimensional}
\newacronym{ram}{RAM}{Random-Access Memory}
\newacronym{cpu}{CPU}{Central Processing Unit}
\newacronym{adsm}{ADSM}{Asymmetric Distributed Shared Memory}
\newacronym{loc}{LOC}{Lines Of Code}
\newacronym{csr}{CSR}{Compressed Storage by Rows}
\newacronym{coo}{COO}{COOrdinate}
\newacronym{csc}{CSC}{Compressed Sparse Columns}
\newacronym{ell}{ELL}{ELLpack}
\newacronym{jad}{JAD}{JAgged Diagonals}
\newacronym{cuda}{CUDA}{Compute Unified Device Architecture}
\newacronym{gflops}{GFLOPS}{Giga (billions) of Floating-point Operations Per Second, or GigaFLOPS}
\newacronym{pde}{PDE}{Partial Differential Equation}
\newacronym{spmv}{SpMV}{Sparse Matrix-Vector multiplication}
\newacronym{ecc}{ECC}{Error-Correcting Code}
\newacronym{petsc}{PETSc}{the Portable Extensible Toolkit for Scientific Computation}
\newacronym{flame}{FLAME}{the Formal Linear Algebra Methods Environment}
\newacronym{cfd}{CFD}{computational fluid dynamics}