### 2017-10-05 response.txt
response.txt list edits made for reviewer comments

Attached please find a revised version of our extended abstract accepted for presentation at PAW17.

Reviewer 1
----------
1. Comment: "[2] Please cite the OpenSHMEM 1.3 specification (Available at: http://openshmem.org)"
   Response: The specification is now reference 10.

2. Comment: "Figure 1: The animation is probably not a good idea for the paper. Why not show a few frames instead?"
   Response: Figure 1 now depicts one representative frame. Space does not permit multiple frames.

3. Comment: "Figures 2-: The 'X' marks are indistinguishable when printed. Please use different symbols.
            Please also show error bars rather than plotting over the same point multiple times."
   Response: The symbols have been changed to better differentiate distinct runs.  We did not perform a sufficient
             number of runs to estimate error bars and the different data points are in fact distinct.  For example,
             the close correspondence between the MPI_put and OSH_put points in Figure 2 stems results from separate
             runs in each case and likely relates to commonalities in the underlying hardware or in the MPI and
             OpenSHMEM implementations.

   Comment: "What about the Intel Fortran compiler? Seems like this compiler should be used on KNL, and it should
            support coarrays?"
   Response: Section 1.2 now details our experiences with the Intel compiler, which produces executable programs
             that crash or hang beyond small core counts.

Review 2
--------

1. Comment (a): "The animation in Figure 1 depicts atmospheric processes over North 105 America as simulated by ICAR".
                Which animation? I would suggest take out animation (whatever that is) and put just a figure or snapshots.
   Response: Figure 1 now depicts one representative frame. Space does not permit multiple frames.

2. Comment (b): "Gets: an image that needs data proactively gets the data from memory managed by a providing image
                without the providing image’s active involvement,": what is an image?
                I know what it is, but it is not defined anywhere before this. A reader that knows nothing about CAF
                should at least be able to read what images are.
   Response: The first paragraph in Section 1 now gives a brief overview of the CAF programming model, including the meaning of "image."

3. Comment (c): "Figure 3 (top) depicts results across multiple compilers, machines, and architectures.
                In this case, only “puts” are tested, and only for the 2000x2000 domain.": why only puts? need to justify
                this more.
4. Response: The second paragraph of Section 3 now explains the reasons only "puts" are shown for MPI.

   Comment (d): in figure 3, lower right: title says runtime comparison, but there is no comparison.
   Response: The word "comparison" has now been deleted.

5. Comment (e): line 392 "This work as funded in..." typo.  "This work was funded in"
   Response: This typographical error has been corrected.

Review 3
--------
Comment: "1.2 spelling of "alteranative" needs to be alternative."
Response: The spelling has been corrected.

Comment: "2.2 The sentence 'We compiled at NCAR using of the GCC...." is poorly worded. Restructure the whole sentence
Response: The cited sentence has been resphrased for greater clarity.